---
title: "COMETmissingness"
author: "Lucy Grigoroff"
date: "2023-11-09"
output: html_document
---
#Complete cases dataset
control is defined by animal numers 1-10 (dose group 1) or timepoint 0
low dose is animal numbers 11-20 (dose group 2) and not timepoint 0
high dose is animal numers 21-30 (dose group 3) and timepoint not 0
```{r cars}
 library(missForest)
 library(mice)
 library(CALIBERrfimpute)
 library(randomForest)
 library(missMethods)
 library(lpSolve)
# devtools::install_github("strengejacke/sjmisc")
 library(sjmisc)
 library(tidyverse)
 library(ggpubr)
library(reshape2)
library(ggpmisc)
library(egg)
library(devEMF)
library(see)
load("/Users/lucygrigoroff/Downloads/new_clinchem_2SD_removed_edited.rda")
df <- new_clinchem

#set up doseFlag column with control, low, high

df <- df %>% mutate(doseFlag =
                     case_when((Dose_group == 1 | Timepoint == 0) ~ "Control",
                               (Dose_group == 2 & Timepoint != 0) ~ "low",
                              (Dose_group == 3 & Timepoint != 0) ~ "high")
)


#serum data frame
#SERUM only has 3 timepoints, 24, 48 and 168
Sdf<- df[,c(1:5,10:13,20:30,32,33)]
idx <- which(Sdf$Timepoint == 24 | df$Timepoint == 48 |df$Timepoint == 168)
Sdf <- Sdf[idx,]
#complete cases
idx<- which(rowSums(is.na(Sdf[10:20])) > 0)
SCdf<-Sdf[-idx,]

SCdf <-SCdf%>%relocate("doseFlag", .before = "S_Creatinine" ) %>% relocate("S_Comments", .before = "S_Creatinine")

# #URINE data frame
# Udf<-df[,c(1:5,10:18,31,33)]
# #complete cases
# idx<- which(rowSums(is.na(Udf[11:14])) > 0)
# UCdf<-Udf[-idx,]
#
# #NA summary
# #SERUM
# colSums(is.na(Sdf))
# as.data.frame(sapply(Sdf[10:20], function(x) sum(is.na(x))/8460*100))
#
# #URINE
# colSums(is.na(Udf))
# as.data.frame(sapply(Udf[11:14], function(x) sum(is.na(x))/25350*100))
```

#choose data size
full, 500 or 10
```{r}

# unique(SCdf$Company)
# random_sample <- data[sample(nrow(data), 50), ]


R <- which(SCdf$Company == "R")
R<- SCdf[R,]
R <- R[sample(nrow(R), 10), ]

D <- which(SCdf$Company == "D")
D <- SCdf[D,]
D <- D[sample(nrow(D), 10), ]

L <- which(SCdf$Company == "L")
L <- SCdf[L,]
L <- L[sample(nrow(L), 10), ]

N <- which(SCdf$Company == "N")
N <- SCdf[N,]
N <- N[sample(nrow(N), 10), ]

S <- which(SCdf$Company == "S")
S <- SCdf[S,]
S <- S[sample(nrow(S), 10), ]

SCdf<-rbind(R, D, L, N, S)

```
#choose columns for imputation input
```{r}
impcols<-c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup")
 #x <- SCdf[,tt]

#to not create missinginess in the following
miscols <- c(1:11)

#column to begin creating missingness
start <- which(colnames(SCdf) == "S_Creatinine")
#or start<-12
end<-ncol(SCdf)

# firstcol = which(colnames(x)=="a")
# lastcol = which(colnames(x)=="b")
#
# mydf[c(firstcol:lastcol)]
# c(start:end)
```

#set up percentages of missingness
```{r}
p <- (cbind(0.05, 0.1, 0.2, 0.3, 0.4))

#for graphs in METRICS and FIGURES
cat(paste(shQuote(paste0( p[1,] * 100, " %"), type="cmd"), collapse=", "))
rn<-c("5 %", "10 %", "20 %", "30 %", "40 %")
```

#list of colnames for MCAR missingess
```{r}
MCARcols <- c("S_Creatinine", "S_Urea_nitrogen", "S_ALT", "S_AST", "S_Glucose","S_Sodium", "S_Potassium", "S_Calcium", "S_Phosphorus", "S_Albumin", "S_Total_protein")

#cat(paste(shQuote(colnames(SCdf), type="cmd"), collapse=", "))

cnames<-c("Company", "Study_id", "Toxin", "Target", "Key", "Animal_no.", "Timepoint", "Sacrifice_subgroup", "Dose_group", "doseFlag", "S_Comments", "S_Creatinine", "S_Urea_nitrogen", "S_ALT", "S_AST", "S_Glucose", "S_Sodium", "S_Potassium", "S_Calcium", "S_Phosphorus", "S_Albumin", "S_Total_protein")

```

#set up original Albumin and Total Protein Bias
```{r}
#set up external checks
original.lm.1 <- summary(lm(S_Albumin ~ S_Total_protein, data = SCdf))
# original.lm.1[["r.squared"]]
original <-original.lm.1[["coefficients"]][2,1]

missed<-list()
missedSummary<-list()
missedBias<-list()
for(j in 1:length(t)){
  missedSummary[[j]] <- summary(lm(S_Albumin ~ S_Total_protein, data = t[[j]], na.action = na.omit))
  missed[[j]] <-missedSummary[[j]][["coefficients"]][2,1]
  missedBias[[j]] <- abs((original - missed[[j]])/original)
}
```
#set up original AST adn ALT
```{r}
####################
original.lm.2 <- summary(lm(S_ALT ~ S_AST, data = SCdf))
# original.lm.1[["r.squared"]]
original2 <-original.lm.2[["coefficients"]][2,1]

missed2<-list()
missedSummary2<-list()
missedBias2<-list()
for(j in 1:length(t)){
  missedSummary2[[j]] <- summary(lm(S_ALT ~ S_AST, data = t[[j]], na.action = na.omit))
  missed2[[j]] <-missedSummary2[[j]][["coefficients"]][2,1]
  missedBias2[[j]] <- abs((original2 - missed2[[j]])/original2)
}
```
dose groups, anything that’s in the control group (animal no. 1-10 or timepoint 0 = 1, animal no 11-20 and not timepoint 0 =2, animal no 20-30 and not timepoint 0=3)

control <- df%>%filter(Dose_group == 1 | Timepoint == 0)
low_dose <- df%>%filter(Dose_group == 2 & Timepoint != 0, )
high_dose <- df%>%filter(Dose_group == 3 & Timepoint != 0, )

MAR
Company and comments "IS=insufficient volume", "IS=Insufficient sample" "No internal code;  IS=Insufficient sample" "No internal code;  IS=Insufficient sample; NC=Not calculated"

missForest: control+low+high: clinchem_params + doseFlag + animal_no + timepoint + sacrifice + company


#MISSINGNESS
Using complete cases for serum, create missingness at varying percentages (5, 10, 20, 30 40)
```{r}


#MAR
MAR <- list()
for (j in (p)) {
  placeHolder <- paste0("missing ", j * 100, " %")

  mSdf <- list()
  test <- SCdf[, miscols]
  for (i in start:ncol(SCdf)) {
    #missingness
    mSdf[[i]] <-
      missMethods::delete_MAR_1_to_x(ds = SCdf,
                                      p = j,
                                      cols_mis = c(i),
                                      cols_ctrl = c("Company"),
                                      x = 3)
    test[, i] <- (mSdf[[i]][[i]])
  }
  new_list <- setNames(list(test), placeHolder)
  MAR <- append(MAR, new_list)
}

for(j in 1:length(MAR)){
  MAR[[j]] <- setNames(MAR[[j]], cnames)
}

#MNAR
MNAR <- list()
for (j in (p)) {
  placeHolder <- paste0("missing ", j * 100, " %")

  mSdf <- list()
  test <- SCdf[, miscols]
  for (i in start:ncol(SCdf)) {
    #missingness
    mSdf[[i]] <-
      missMethods::delete_MNAR_1_to_x(ds = SCdf,
                                      p = j,
                                      cols_mis = c(i),
                                      x = 3)
    test[, i] <- (mSdf[[i]][[i]])
  }
  new_list <- setNames(list(test), placeHolder)
  MNAR <- append(MNAR, new_list)
}

for(j in 1:length(MNAR)){
  MNAR[[j]] <- setNames(MNAR[[j]], cnames)
}

#MCAR
MCAR<-list()
for(j in p) {
  placeHolder <- paste0("missing ", j * 100, " %")
  mSdf <- list()
  #missingness
  mSdf[[i]] <- missMethods::delete_MCAR(ds = SCdf, p = j, MCARcols)
  new_list <- setNames(list(mSdf[[i]]), placeHolder)
  MCAR <- append(MCAR, new_list)

}

rm(mSdf, new_list, test, new_clinchem,Sdf, df)
```
