---
title: "imputation"
author: "Lucy Grigoroff"
date: "2023-08-28"
output: html_document
---
```{r}
set.seed(1)
```

#MISSINGNESS
##Complete cases dataset
control is defined by animal numers 1-10 (dose group 1) or timepoint 0
low dose is animal numbers 11-20 (dose group 2) and not timepoint 0
high dose is animal numers 21-30 (dose group 3) and timepoint not 0
```{r cars}
 library(missForest)
 library(mice)
 library(CALIBERrfimpute)
 library(randomForest)
 library(missMethods)
 library(lpSolve)
# devtools::install_github("strengejacke/sjmisc")
 library(sjmisc)
 library(tidyverse)
 library(ggpubr)
library(reshape2)
library(ggpmisc)
library(egg)
library(devEMF)
library(see)

load("~/OneDrive - Murdoch University/COMET/COMET/new_clinchem_2SD_removed_edited.rda")

df <- new_clinchem

#set up doseFlag column with control, low, high

df <- df %>% mutate(doseFlag =
                     case_when((Dose_group == 1 | Timepoint == 0) ~ "Control",
                               (Dose_group == 2 & Timepoint != 0) ~ "low",
                              (Dose_group == 3 & Timepoint != 0) ~ "high")
)


#serum data frame
#SERUM only has 3 timepoints, 24, 48 and 168
Sdf<- df[,c(1:5,10:13,20:30,32,33)]
idx <- which(Sdf$Timepoint == 24 | df$Timepoint == 48 |df$Timepoint == 168)
Sdf <- Sdf[idx,]
#complete cases
idx<- which(rowSums(is.na(Sdf[10:20])) > 0)
SCdf<-Sdf[-idx,]

SCdf <-SCdf%>%relocate("doseFlag", .before = "S_Creatinine" ) %>% relocate("S_Comments", .before = "S_Creatinine")

# #URINE data frame
# Udf<-df[,c(1:5,10:18,31,33)]
# #complete cases
# idx<- which(rowSums(is.na(Udf[11:14])) > 0)
# UCdf<-Udf[-idx,]
#
# #NA summary
# #SERUM
# colSums(is.na(Sdf))
# as.data.frame(sapply(Sdf[10:20], function(x) sum(is.na(x))/8460*100))
#
# #URINE
# colSums(is.na(Udf))
# as.data.frame(sapply(Udf[11:14], function(x) sum(is.na(x))/25350*100))
```

##choose data size
full, 500 or 10
```{r}

# unique(SCdf$Company)
# random_sample <- data[sample(nrow(data), 50), ]


R <- which(SCdf$Company == "R")
R<- SCdf[R,]
R <- R[sample(nrow(R), 100), ]

D <- which(SCdf$Company == "D")
D <- SCdf[D,]
D <- D[sample(nrow(D), 100), ]

L <- which(SCdf$Company == "L")
L <- SCdf[L,]
L <- L[sample(nrow(L), 100), ]

N <- which(SCdf$Company == "N")
N <- SCdf[N,]
N <- N[sample(nrow(N), 100), ]

S <- which(SCdf$Company == "S")
S <- SCdf[S,]
S <- S[sample(nrow(S), 100), ]

SCdf<-rbind(R, D, L, N, S)

SCdf500full <- SCdf

```


##set up percentages of missingness
```{r}
p <- (cbind(0.05, 0.1, 0.2, 0.3, 0.4))

#for graphs in METRICS and FIGURES
cat(paste(shQuote(paste0( p[1,] * 100, " %"), type="cmd"), collapse=", "))
rn<-c("5 %", "10 %", "20 %", "30 %", "40 %")
```

##list of colnames for MCAR missingess
```{r}
MCARcols <- c("S_Creatinine", "S_Urea_nitrogen", "S_ALT", "S_AST", "S_Glucose","S_Sodium", "S_Potassium", "S_Calcium", "S_Phosphorus", "S_Albumin", "S_Total_protein")

#cat(paste(shQuote(colnames(SCdf), type="cmd"), collapse=", "))

cnames<-c("Company", "Study_id", "Toxin", "Target", "Key", "Animal_no.", "Timepoint", "Sacrifice_subgroup", "Dose_group", "doseFlag", "S_Comments", "S_Creatinine", "S_Urea_nitrogen", "S_ALT", "S_AST", "S_Glucose", "S_Sodium", "S_Potassium", "S_Calcium", "S_Phosphorus", "S_Albumin", "S_Total_protein")

```

dose groups, anything that’s in the control group (animal no. 1-10 or timepoint 0 = 1, animal no 11-20 and not timepoint 0 =2, animal no 20-30 and not timepoint 0=3)

control <- df%>%filter(Dose_group == 1 | Timepoint == 0)
low_dose <- df%>%filter(Dose_group == 2 & Timepoint != 0, )
high_dose <- df%>%filter(Dose_group == 3 & Timepoint != 0, )

MAR
Company and comments "IS=insufficient volume", "IS=Insufficient sample" "No internal code;  IS=Insufficient sample" "No internal code;  IS=Insufficient sample; NC=Not calculated"

missForest: control+low+high: clinchem_params + doseFlag + animal_no + timepoint + sacrifice + company


##set up missingness 
Using complete cases for serum, create missingness at varying percentages (5, 10, 20, 30 40)
```{r}
#to not create missinginess in the following
miscols <- c(1:11)

#MAR
MAR <- list()
for (j in (p)) {
  placeHolder <- paste0("missing ", j * 100, " %")

  mSdf <- list()
  test <- SCdf[, miscols]
  for (i in start:ncol(SCdf)) {
    #missingness
    mSdf[[i]] <-
      missMethods::delete_MAR_1_to_x(ds = SCdf,
                                      p = j,
                                      cols_mis = c(i),
                                      cols_ctrl = c("Company"),
                                      x = 3)
    test[, i] <- (mSdf[[i]][[i]])
  }
  new_list <- setNames(list(test), placeHolder)
  MAR <- append(MAR, new_list)
}

for(j in 1:length(MAR)){
  MAR[[j]] <- setNames(MAR[[j]], cnames)
}

#MNAR
MNAR <- list()
for (j in (p)) {
  placeHolder <- paste0("missing ", j * 100, " %")

  mSdf <- list()
  test <- SCdf[, miscols]
  for (i in start:ncol(SCdf)) {
    #missingness
    mSdf[[i]] <-
      missMethods::delete_MNAR_1_to_x(ds = SCdf,
                                      p = j,
                                      cols_mis = c(i),
                                      x = 3)
    test[, i] <- (mSdf[[i]][[i]])
  }
  new_list <- setNames(list(test), placeHolder)
  MNAR <- append(MNAR, new_list)
}

for(j in 1:length(MNAR)){
  MNAR[[j]] <- setNames(MNAR[[j]], cnames)
}

#MCAR
MCAR<-list()
for(j in p) {
  placeHolder <- paste0("missing ", j * 100, " %")
  mSdf <- list()
  #missingness
  mSdf[[i]] <- missMethods::delete_MCAR(ds = SCdf, p = j, MCARcols)
  new_list <- setNames(list(mSdf[[i]]), placeHolder)
  MCAR <- append(MCAR, new_list)

}

rm(mSdf, new_list, test, new_clinchem,Sdf, df)

#if need to save
# MCAR500<-MCAR
# MAR500<-MAR
# MNAR500<-MNAR
```

#IMPUTAION
##choose columns for imputation input
```{r}
#for 500 samples
#load("~/OneDrive - Murdoch University/COMET/COMET/500 Samples/500.Rdata")
#SCdf <- SCdf500full

#for 50 samples
#load("~/OneDrive - Murdoch University/COMET/COMET/50 Samples/50.Rdata")
#SCdf <- SCdf50full 
 

#which columns are included in the imputation 
impcols<-
  #NULL
  #"Company"
  #c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup", "Target")
  #c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup", "Target", "Toxin")

#column to begin creating missingness
start <- which(colnames(SCdf) == "S_Creatinine")
#or start<-12
end<-ncol(SCdf)

```

##choose missingess method
```{r}
t<- MCAR
full<-SCdf[,start:end]
#full500<-full

```

##set up original Albumin and Total Protein Bias
DO FOR EACH MISSING METHOD (MCAR, MAR, MNAR)
```{r}
#set up external checks
original.lm.1 <- summary(lm(S_Albumin ~ S_Total_protein, data = SCdf))
# original.lm.1[["r.squared"]]
original <-original.lm.1[["coefficients"]][2,1]

missed<-list()
missedSummary<-list()
missedBias<-list()
for(j in 1:length(t)){
  missedSummary[[j]] <- summary(lm(S_Albumin ~ S_Total_protein, data = t[[j]], na.action = na.omit))
  missed[[j]] <-missedSummary[[j]][["coefficients"]][2,1]
  missedBias[[j]] <- abs((original - missed[[j]])/original)

}
```
##set up original AST adn ALT
DO FOR EACH MISSING METHOD (MCAR, MAR, MNAR)
```{r}
####################
original.lm.2 <- summary(lm(S_ALT ~ S_AST, data = SCdf))
# original.lm.1[["r.squared"]]
original2 <-original.lm.2[["coefficients"]][2,1]

missed2<-list()
missedSummary2<-list()
missedBias2<-list()
for(j in 1:length(t)){
  missedSummary2[[j]] <- summary(lm(S_ALT ~ S_AST, data = t[[j]], na.action = na.omit))
  missed2[[j]] <-missedSummary2[[j]][["coefficients"]][2,1]
  missedBias2[[j]] <- abs((original2 - missed2[[j]])/original2)
}

```




##MICErf
DO FOR EACH MISSING METHOD (MCAR, MAR, MNAR)
For the following column combinations:
  NULL
  "Company"
  c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup", "Target")
  c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup", "Target", "Toxin")
```{r}

 res<-list()
for(k in 1:20){
#full<-SCdf[,12:22]
test<-list()
imp<-list()
imp2<-list()
imputedonly<-list()
MICErf<-list()
FULL<-list()
PARTIAL<-list()
for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  
  if(length(impcols) == 1){
   x<-as.data.frame(SCdf[,impcols])
   colnames(x) <- c("Company")
 }

  if(length(impcols) > 1){
   x<-SCdf[,impcols]
  }

   if(length(impcols) < 1){
   x<-as.data.frame(rep_len(x = 1, length.out = nrow(SCdf)))
 }
  
  #make predictors numeric

 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)
 
#make the predictor matrix using all covariates (including clin_chem parameters)
x<-cbind(x,t[[j]][,start:end])
init = mice(x, maxit=0) 
predM = init$predictorMatrix

#Do imputation
imp[[j]]<-mice(data = x, method="rf", predictorMatrix = predM)
imp2[[j]]<-merge_imputations(dat = x, imp = imp[[j]])
idx<- as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))
idx$imputed<-idx$original<-NA

for(i in 1:nrow(idx)){
  idx$original[i]<-unlist(full[idx$row[i],idx$col[i]])
  idx$imputed[i]<-imp2[[j]][idx$row[i],idx$col[i]]
  imputedonly[[j]]<-idx
}

new_list <- setNames(list(imp2[[j]]), placeHolder)
FULL <- append(FULL, new_list)  

new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
PARTIAL <- append(PARTIAL, new_list2) 
}

MICErf<- append(MICErf, list(PARTIAL = PARTIAL,
                             FULL = FULL))

#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(MICErf[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(MICErf[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(MICErf[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- MICErf[["PARTIAL"]][[j]][["original"]]
  xi <- MICErf[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
#External check 1 MICErf
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.1 <- list()
for(j in 1:length(MICErf[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
  imputedSummary[[j]] <- summary(lm(S_Albumin ~ S_Total_protein, data = MICErf[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original - imputed[[j]])/original)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.1 <- append(imputedBias.1, new_list) 
}

#External Check 2 MICErf
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.2 <- list()
for(j in 1:length(MICErf[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
  imputedSummary[[j]] <- summary(lm(S_ALT ~ S_AST, data = MICErf[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original2 - imputed[[j]])/original2)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.2 <- append(imputedBias.2, new_list) 
}


MICErf<- append(MICErf, list(NRMSE = NRMSE,
                             partialMAE = partialMAE,
                             fullMAE = fullMAE,
                             imputedBias.1 = imputedBias.1,
                             imputedBias.2 = imputedBias.2))


res[[k]] <- MICErf
}

#nameing

 if((MCAR[1] %in% t[1])){
   missing<-"_MCAR"
 }

if((MAR[1] %in% t[1]) ){
  missing <-"_MAR"
}

if((MNAR[1] %in% t[1]) ){
  missing <-"_MNAR"
}

if(length(impcols) == 1){
   impco<- "_Company"
 }

if(length(impcols) > 1 && !("Toxin" %in% impcols)){
  impco<- "_All"
  }

if(length(impcols) > 1 && "Toxin" %in% impcols){
  impco<- "_All_Toxin"
  }

 if(length(impcols) < 1){
   impco<-"_noMeta"
 }

list_name <- paste("MICErf", missing, impco, sep = "")
assign(list_name, res, envir = .GlobalEnv)


n_entries <- length(res)
metrics <- c("NRMSE", "partialMAE", "fullMAE", "imputedBias.1", "imputedBias.2")

# Initialize a list to store the results
results <- list()

# Loop through the metrics
for (metric in metrics) {
  result_matrix <- matrix(NA, nrow = n_entries, ncol = length(res[[1]][[metric]]))
  
  # Loop through the list and extract the current metric values into the matrix
  for (i in 1:n_entries) {
    result_matrix[i, ] <- unlist(res[[i]][[metric]])
  }
  
  # Calculate mean and standard deviation for each column
  mean_values <- apply(result_matrix, 2, mean)
  std_dev_values <- apply(result_matrix, 2, sd)
  
  results[[metric]] <- list(mean = mean_values, std_dev = std_dev_values)
}

list_name <- paste("summary_MICErf", missing, impco, sep = "")
assign(list_name, results, envir = .GlobalEnv)



```

##MICErf strata
Repeat the below process for MCAR, MAR and MNAR
For the following column combinations:
  "Company"
  c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup", "Target")
```{r}


  res<-list()
for(k in 1:20){
#full<-SCdf[,12:22]
test<-list()
imp<-list()
imp2<-list()
imputedonly<-list()
MICErf<-list()
FULL<-list()
PARTIAL<-list()
for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  
    if(length(impcols) == 1){
   x<-as.data.frame(SCdf[,impcols])
   colnames(x) <- c("Company")
 }

  if(length(impcols) > 1){
   x<-SCdf[,impcols]
  }

   if(length(impcols) < 1){
   x<-as.data.frame(rep_len(x = 1, length.out = nrow(SCdf)))
 }
  
  #make predictors numeric

 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)
 
#make the predictor matrix using all covariates (including clin_chem parameters)
x<-cbind(x,t[[j]][,start:end])
init = mice(x, maxit=0) 
predM = init$predictorMatrix

#Do imputation
# imp[[j]]<-mice(data = x, method="rf", predictorMatrix = predM)
# imp2[[j]]<-merge_imputations(dat = x, imp = imp[[j]])


# #Do imputation
# 
stratimp2<- list()
stratimp<-list()
for(i in 1:5){
stratimp<-mice(data = x[which(x$Company == i),], method="rf", predictorMatrix = predM)
stratimp2[[i]]<-merge_imputations(dat = x[which(x$Company == i),], imp = stratimp)

#if the number of columns are less than 11, then one of them did not contain an NA post stratification, which will prevent the next step. If this is the case, find the column, reattach is and put it in the correct/same order as the others. May only work when it's one column missing

if(ncol(stratimp2[[i]]) < 11){
  #find missing column
base <- c("S_Creatinine", "S_Urea_nitrogen", "S_ALT", "S_AST", "S_Glucose", "S_Sodium", "S_Potassium", "S_Calcium", "S_Phosphorus", "S_Albumin", "S_Total_protein")
missing_cols <- setdiff(base, colnames(stratimp2[[i]]))

#reattach the column
stratimp2[[i]][,missing_cols] <- x[which(x$Company == i), missing_cols]

# Reorder columns in the data frame
stratimp2[[i]] <- stratimp2[[i]][, base]
}

}

#must stack the companies in the correct order ("R" "D" "L" "N" "S" so 4, 1, 2, 3, 5)
test <- rbind(stratimp2[[4]], stratimp2[[1]], stratimp2[[2]], stratimp2[[3]], stratimp2[[5]])

imp2[[j]] <- test

 idx<- as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))
 idx$imputed<-idx$original<-NA
for(i in 1:nrow(idx)){
  idx$original[i]<-unlist(full[idx$row[i],idx$col[i]])
  idx$imputed[i]<-imp2[[j]][idx$row[i],idx$col[i]]
  imputedonly[[j]]<-idx
}

new_list <- setNames(list(imp2[[j]]), placeHolder)
FULL <- append(FULL, new_list)  

new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
PARTIAL <- append(PARTIAL, new_list2) 
}

MICErf<- append(MICErf, list(PARTIAL = PARTIAL,
                             FULL = FULL))

#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(MICErf[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(MICErf[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(MICErf[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- MICErf[["PARTIAL"]][[j]][["original"]]
  xi <- MICErf[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
#External check 1 MICErf
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.1 <- list()
for(j in 1:length(MICErf[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
  imputedSummary[[j]] <- summary(lm(S_Albumin ~ S_Total_protein, data = MICErf[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original - imputed[[j]])/original)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.1 <- append(imputedBias.1, new_list) 
}

#External Check 2 MICErf
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.2 <- list()
for(j in 1:length(MICErf[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
  imputedSummary[[j]] <- summary(lm(S_ALT ~ S_AST, data = MICErf[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original2 - imputed[[j]])/original2)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.2 <- append(imputedBias.2, new_list) 
}


MICErf<- append(MICErf, list(NRMSE = NRMSE,
                             partialMAE = partialMAE,
                             fullMAE = fullMAE,
                             imputedBias.1 = imputedBias.1,
                             imputedBias.2 = imputedBias.2))


res[[k]] <- MICErf
}

#nameing

 if((MCAR[1] %in% t[1])){
   missing<-"_MCAR"
 }

if((MAR[1] %in% t[1]) ){
  missing <-"_MAR"
}

if((MNAR[1] %in% t[1]) ){
  missing <-"_MNAR"
}

if(length(impcols) == 1){
   impco<- "_Company"
 }

if(length(impcols) > 1 && !("Toxin" %in% impcols)){
  impco<- "_All"
  }

if(length(impcols) > 1 && "Toxin" %in% impcols){
  impco<- "_All_Toxin"
  }

 if(length(impcols) < 1){
   impco<-"_noMeta"
 }

list_name <- paste("MICErf_strata", missing, impco, sep = "")
assign(list_name, res, envir = .GlobalEnv)


n_entries <- length(res)
metrics <- c("NRMSE", "partialMAE", "fullMAE", "imputedBias.1", "imputedBias.2")

# Initialize a list to store the results
results <- list()

# Loop through the metrics
for (metric in metrics) {
  result_matrix <- matrix(NA, nrow = n_entries, ncol = length(res[[1]][[metric]]))
  
  # Loop through the list and extract the current metric values into the matrix
  for (i in 1:n_entries) {
    result_matrix[i, ] <- unlist(res[[i]][[metric]])
  }
  
  # Calculate mean and standard deviation for each column
  mean_values <- apply(result_matrix, 2, mean)
  std_dev_values <- apply(result_matrix, 2, sd)
  
  results[[metric]] <- list(mean = mean_values, std_dev = std_dev_values)
}

list_name <- paste("summary_MICErf_strata", missing, impco, sep = "")
assign(list_name, results, envir = .GlobalEnv)



```



##missForest
Repeat the below process for MCAR, MAR and MNAR
For the following column combinations:
  NULL
  "Company"
  c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup", "Target")
  
You cannot include Toxin as missForest only allows a max of 52 levels for a factor, Toxin has more than this. 
```{r}


res<-list()
for(k in 1:20){
  
FULL<-list()
PARTIAL <- list()
missForest<-list()
imp <- list()
imputedonly<-list()
nrmse<-list()
NRMSE<-list()
for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  placeHolder3<- paste0("NRMSE ", p[j] * 100, " % missing")
  
if(length(impcols) == 1){
   x<-as.data.frame(SCdf[,impcols])
   colnames(x) <- c("Company")
 }

  if(length(impcols) > 1){
   x<-SCdf[,impcols]
  }

   if(length(impcols) < 1){
   x<-as.data.frame(rep_len(x = 1, length.out = nrow(SCdf)))
 }  
  
  #make predictors factors (can also be numeric, just not characters, max of 53 categories per variable allowed)
 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 
 #set up data frame used in imputation 
 x <- cbind(x,t[[1]][,start:end])
 
 #impute
 imp[[j]] <- missForest(xmis = x, ntree = 10, variablewise = F)
 
 #locate where in the complete data the NAs were created
idx<- as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))

#make empty list to fill with original and imputed
idx$imputed<-idx$original<-NA

#start and end criteria 
start2 <- which(colnames(x) == "S_Creatinine")
end2 <- ncol(x)

for(i in 1:nrow(idx)){
  
  #fill original column
  idx$original[i]<-full[idx$row[i],idx$col[i]]
  #fill imputed column
  idx$imputed[i]<-imp[[j]][["ximp"]][,start2:end2][idx$row[i],idx$col[i]]
  #store imputed only for each percentage of missingness
  imputedonly[[j]]<-idx
}

new_list <- setNames(list(imp[[j]][["ximp"]][,start2:end2]), placeHolder)
FULL <- append(FULL, new_list)  

 new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
 PARTIAL <- append(PARTIAL, new_list2) 

}

missForest<- append(missForest, list(PARTIAL = PARTIAL,
                                     FULL = FULL
                                     ))

#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(missForest[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(missForest[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(missForest[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- missForest[["PARTIAL"]][[j]][["original"]]
  xi <- missForest[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
#External check 1 missForest
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.1 <- list()
for(j in 1:length(missForest[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")
  imputedSummary[[j]] <- summary(lm(S_Albumin ~ S_Total_protein, data = missForest[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original - imputed[[j]])/original)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.1 <- append(imputedBias.1, new_list)
}

#External check 2 missForest
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.2 <- list()
for(j in 1:length(missForest[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")
  imputedSummary[[j]] <- summary(lm(S_ALT ~ S_AST, data = missForest[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original2 - imputed[[j]])/original2)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.2 <- append(imputedBias.2, new_list)
}
 
missForest<- append(missForest, list(NRMSE =NRMSE,
                                     partialMAE = partialMAE,
                                     fullMAE = fullMAE ,
                                      imputedBias.1 = imputedBias.1,
                                      imputedBias.2 = imputedBias.2
                                     ))  

res[[k]] <- missForest
}
#nameing

 if((MCAR[1] %in% t[1])){
   missing<-"_MCAR"
 }

if((MAR[1] %in% t[1]) ){
  missing <-"_MAR"
}

if((MNAR[1] %in% t[1]) ){
  missing <-"_MNAR"
}

if(length(impcols) == 1){
   impco<- "_Company"
 }

if(length(impcols) > 1 && !("Toxin" %in% impcols)){
  impco<- "_All"
  }

if(length(impcols) > 1 && "Toxin" %in% impcols){
  impco<- "_All_Toxin"
  }

 if(length(impcols) < 1){
   impco<-"_noMeta"
 }

list_name <- paste("missForest", missing, impco, sep = "")
assign(list_name, res, envir = .GlobalEnv)

n_entries <- length(res)
metrics <- c("NRMSE", "partialMAE", "fullMAE", "imputedBias.1", "imputedBias.2")

# Initialize a list to store the results
results <- list()

# Loop through the metrics
for (metric in metrics) {
  result_matrix <- matrix(NA, nrow = n_entries, ncol = length(res[[1]][[metric]]))
  
  # Loop through the list and extract the current metric values into the matrix
  for (i in 1:n_entries) {
    result_matrix[i, ] <- unlist(res[[i]][[metric]])
  }
  
  # Calculate mean and standard deviation for each column
  mean_values <- apply(result_matrix, 2, mean)
  std_dev_values <- apply(result_matrix, 2, sd)
  
  results[[metric]] <- list(mean = mean_values, std_dev = std_dev_values)
}

list_name <- paste("summary_missForest", missing, impco, sep = "")
assign(list_name, results, envir = .GlobalEnv)

```

##missForest strata
Repeat the below process for MCAR, MAR and MNAR
For the following column combinations:
  "Company"
  c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup", "Target")
```{r}
res<-list()
for(k in 1:20){
  
FULL<-list()
PARTIAL <- list()
missForest<-list()
imp <- list()
imputedonly<-list()
nrmse<-list()
NRMSE<-list()
for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  placeHolder3<- paste0("NRMSE ", p[j] * 100, " % missing")
  
  
  if(length(impcols) == 1){
   x<-as.data.frame(SCdf[,impcols])
   colnames(x) <- c("Company")
 }

  if(length(impcols) > 1){
   x<-SCdf[,impcols]
  }

   if(length(impcols) < 1){
   x<-as.data.frame(rep_len(x = 1, length.out = nrow(SCdf)))
 }
  #make predictors factors (can also be numeric, just not characters, max of 53 categories per variable allowed)

 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 
 #set up data frame used in imputation 
 x <- cbind(x,t[[1]][,start:end])
 
 #impute
 
  ##################
#rm(stratimp)
stratimp2<- list()
companies <- c("D", "L", "N", "R", "S")

for (z in 1:5) {
  stratimp2[[z]] <- missForest(xmis = x[x$Company == companies[z], ], ntree = 10, variablewise = FALSE)
}
test <- rbind(stratimp2[[1]][["ximp"]], stratimp2[[2]][["ximp"]], stratimp2[[3]][["ximp"]], stratimp2[[4]][["ximp"]], stratimp2[[5]][["ximp"]])

imp[[j]] <- test

 ##################
 #imp[[j]] <- missForest(xmis = x, ntree = 10, variablewise = F)
 
 #locate where in the complete data the NAs were created
idx<- as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))

#make empty list to fill with original and imputed
idx$imputed<-idx$original<-NA

#start and end criteria 
start2 <- which(colnames(x) == "S_Creatinine")
end2 <- ncol(x)

for(i in 1:nrow(idx)){
  
  #fill original column
  idx$original[i]<-full[idx$row[i],idx$col[i]]
  #fill imputed column
  idx$imputed[i]<-imp[[j]][,start2:end2][idx$row[i],idx$col[i]]
  #store imputed only for each percentage of missingness
  imputedonly[[j]]<-idx
}

new_list <- setNames(list(imp[[j]][,start2:end2]), placeHolder)
FULL <- append(FULL, new_list)  

 new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
 PARTIAL <- append(PARTIAL, new_list2) 

}

missForest<- append(missForest, list(PARTIAL = PARTIAL,
                                     FULL = FULL
                                     ))

#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(missForest[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(missForest[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(missForest[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- missForest[["PARTIAL"]][[j]][["original"]]
  xi <- missForest[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
#External check 1 missForest
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.1 <- list()
for(j in 1:length(missForest[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")
  imputedSummary[[j]] <- summary(lm(S_Albumin ~ S_Total_protein, data = missForest[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original - imputed[[j]])/original)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.1 <- append(imputedBias.1, new_list)
}

#External check 2 missForest
imputed<-list()
imputedSummary<-list()
imputedBias<-list()
imputedBias.2 <- list()
for(j in 1:length(missForest[["FULL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")
  imputedSummary[[j]] <- summary(lm(S_ALT ~ S_AST, data = missForest[["FULL"]][[j]]))
  imputed[[j]] <-imputedSummary[[j]][["coefficients"]][2,1]
  imputedBias[[j]]<- abs((original2 - imputed[[j]])/original2)
  new_list <- setNames((imputedBias[[j]]), placeHolder3)
  imputedBias.2 <- append(imputedBias.2, new_list)
}
 
missForest<- append(missForest, list(NRMSE =NRMSE,
                                     partialMAE = partialMAE,
                                     fullMAE = fullMAE ,
                                      imputedBias.1 = imputedBias.1,
                                      imputedBias.2 = imputedBias.2
                                     ))  

res[[k]] <- missForest
}
#nameing

 if((MCAR[1] %in% t[1])){
   missing<-"_MCAR"
 }

if((MAR[1] %in% t[1]) ){
  missing <-"_MAR"
}

if((MNAR[1] %in% t[1]) ){
  missing <-"_MNAR"
}

if(length(impcols) == 1){
   impco<- "_Company"
 }

if(length(impcols) > 1 && !("Toxin" %in% impcols)){
  impco<- "_All"
  }

if(length(impcols) > 1 && "Toxin" %in% impcols){
  impco<- "_All_Toxin"
  }

 if(length(impcols) < 1){
   impco<-"_noMeta"
 }

list_name <- paste("missForest_strata", missing, impco, sep = "")
assign(list_name, res, envir = .GlobalEnv)

n_entries <- length(res)
metrics <- c("NRMSE", "partialMAE", "fullMAE", "imputedBias.1", "imputedBias.2")

# Initialize a list to store the results
results <- list()

# Loop through the metrics
for (metric in metrics) {
  result_matrix <- matrix(NA, nrow = n_entries, ncol = length(res[[1]][[metric]]))
  
  # Loop through the list and extract the current metric values into the matrix
  for (i in 1:n_entries) {
    result_matrix[i, ] <- unlist(res[[i]][[metric]])
  }
  
  # Calculate mean and standard deviation for each column
  mean_values <- apply(result_matrix, 2, mean)
  std_dev_values <- apply(result_matrix, 2, sd)
  
  results[[metric]] <- list(mean = mean_values, std_dev = std_dev_values)
}

list_name <- paste("summary_missForest_strata", missing, impco, sep = "")
assign(list_name, results, envir = .GlobalEnv)

```

###Save

Save after completing MCAR MAR and MNAR for a particular sample size
```{r}
save(
       summary_MICErf_MCAR_All,
summary_MICErf_MCAR_All_Toxin,
       summary_MICErf_MCAR_Company,
       summary_MICErf_MCAR_noMeta,
               summary_MICErf_strata_MCAR_All,
               summary_MICErf_strata_MCAR_Company,
       summary_missForest_MCAR_All,
       summary_missForest_MCAR_Company,
       summary_missForest_MCAR_noMeta,
               summary_missForest_strata_MCAR_All,
               summary_missForest_strata_MCAR_Company,
       MICErf_MCAR_All,
MICErf_MCAR_All_Toxin,
       MICErf_MCAR_Company,
       MICErf_MCAR_noMeta,
               MICErf_strata_MCAR_All,
               MICErf_strata_MCAR_Company,
       missForest_MCAR_All,
       missForest_MCAR_Company,
       missForest_MCAR_noMeta,
               missForest_strata_MCAR_All,
               missForest_strata_MCAR_Company,
     file = "MCAR500iteration.Rdata")

```


