---
title: "imputation"
author: "Lucy Grigoroff"
date: "2023-08-28"
output: html_document
---

#Complete cases dataset
control is defined by animal numers 1-10 (dose group 1) or timepoint 0 
low dose is animal numbers 11-20 (dose group 2) and not timepoint 0 
high dose is animal numers 21-30 (dose group 3) and timepoint not 0
```{r cars}
 library(missForest)
 library(randomForest)
 library(missMethods)
 library(lpSolve)
# devtools::install_github("strengejacke/sjmisc")
 library(sjmisc)
 library(tidyverse)

load("/Users/lucygrigoroff/Downloads/new_clinchem_2SD_removed_edited.rda")
df <- new_clinchem

#set up doseFlag column with control, low, high

df <- df %>% mutate(doseFlag =
                     case_when((Dose_group == 1 | Timepoint == 0) ~ "Control", 
                               (Dose_group == 2 & Timepoint != 0) ~ "low",
                              (Dose_group == 3 & Timepoint != 0) ~ "high")
)


#serum data frame
#SERUM only has 3 timepoints, 24, 48 and 168
Sdf<- df[,c(1:5,10:13,20:30,32,33)]
idx <- which(Sdf$Timepoint == 24 | df$Timepoint == 48 |df$Timepoint == 168)  
Sdf <- Sdf[idx,]
#complete cases
idx<- which(rowSums(is.na(Sdf[10:20])) > 0)
SCdf<-Sdf[-idx,]

SCdf <-SCdf%>%relocate("doseFlag", .before = "S_Creatinine" ) %>% relocate("S_Comments", .before = "S_Creatinine")
# #URINE data frame
# Udf<-df[,c(1:5,10:18,31,33)]
# #complete cases
# idx<- which(rowSums(is.na(Udf[11:14])) > 0)
# UCdf<-Udf[-idx,]
# 
# #NA summary
# #SERUM
# colSums(is.na(Sdf))
# as.data.frame(sapply(Sdf[10:20], function(x) sum(is.na(x))/8460*100))
# 
# #URINE
# colSums(is.na(Udf))
# as.data.frame(sapply(Udf[11:14], function(x) sum(is.na(x))/25350*100))
```

#choose data size 
full, 500 or 10
```{r}

```
#choose columns for imputation input
```{r}
impcols<-c("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup")
 #x <- SCdf[,tt]

#to not create missinginess in the following
miscols <- c(1:11)

#column to begin creating missingness
start <- which(colnames(SCdf) == "S_Creatinine")
#or start<-12
end<-ncol(SCdf)

# firstcol = which(colnames(x)=="a")
# lastcol = which(colnames(x)=="b")
# 
# mydf[c(firstcol:lastcol)]
# c(start:end)
```

#set up percentages of missingness
```{r}
p <- (cbind(0.05, 0.1, 0.2, 0.3, 0.4))
```

#list of colnames for MCAR missingess
```{r}
MCARcols <- c("S_Creatinine", "S_Urea_nitrogen", "S_ALT", "S_AST", "S_Glucose","S_Sodium", "S_Potassium", "S_Calcium", "S_Phosphorus", "S_Albumin", "S_Total_protein")
```
dose groups, anything that’s in the control group (animal no. 1-10 or timepoint 0 = 1, animal no 11-20 and not timepoint 0 =2, animal no 20-30 and not timepoint 0=3)

control <- df%>%filter(Dose_group == 1 | Timepoint == 0)
low_dose <- df%>%filter(Dose_group == 2 & Timepoint != 0, )
high_dose <- df%>%filter(Dose_group == 3 & Timepoint != 0, )

MAR
Company and comments "IS=insufficient volume", "IS=Insufficient sample" "No internal code;  IS=Insufficient sample" "No internal code;  IS=Insufficient sample; NC=Not calculated"   

missForest: control+low+high: clinchem_params + doseFlag + animal_no + timepoint + sacrifice + company
[1] "Company"   "Sacrifice_subgroup"    "Toxin"   "Animal_no."     "Dose_group"       "Timepoint"    

#MISSINGNESS
Using complete cases for serum, create missingness at varying percentages (5, 10, 20, 30 40)
```{r}
#SCdf <-SCdf%>%relocate("doseFlag", .before = "S_Creatinine" ) %>% relocate("S_Comments", .before = "S_Creatinine")
#p <- (cbind(0.05, 0.1, 0.2, 0.3, 0.4))

#MAR
MAR <- list()
for (j in (p)) {
  placeHolder <- paste0("missing ", j * 100, " %")
  
  mSdf <- list()
  test <- SCdf[, miscols]
  for (i in start:ncol(SCdf)) {
    #missingness
    mSdf[[i]] <-
      missMethods::delete_MAR_1_to_x(ds = SCdf,
                                      p = j,
                                      cols_mis = c(i),
                                      cols_ctrl = c("Company"),
                                      x = 3)
    test[, i] <- (mSdf[[i]][[i]])
  }
  new_list <- setNames(list(test), placeHolder)
  MAR <- append(MAR, new_list)
}

#MNAR
MNAR <- list()
for (j in (p)) {
  placeHolder <- paste0("missing ", j * 100, " %")
  
  mSdf <- list()
  test <- SCdf[, miscols]
  for (i in start:ncol(SCdf)) {
    #missingness
    mSdf[[i]] <-
      missMethods::delete_MNAR_1_to_x(ds = SCdf,
                                      p = j,
                                      cols_mis = c(i),
                                      x = 3)
    test[, i] <- (mSdf[[i]][[i]])
  }
  new_list <- setNames(list(test), placeHolder)
  MNAR <- append(MNAR, new_list)
}

#MCAR
MCAR<-list()  
for(j in p){
  placeHolder <- paste0("missing ", j * 100, " %")
  mSdf <-list()
     #missingness
  mSdf[[i]] <- missMethods::delete_MCAR(ds = SCdf, p = j, MCARcols)
  new_list <- setNames(list(mSdf[[i]]), placeHolder)
  MCAR <- append(MCAR, new_list)

}

rm(mSdf, new_list, test, new_clinchem,Sdf, df)
```

#choose missingess method

Need to change the way the eg CALIBER_MAR list is set up. Use paste 
```{r}
t<- MAR
full<-SCdf[,start:end]
```


#IMPUTATION
##CALIBERrfimpute
###CALIBERrfimpute without clinchem

CALIBERrfimpute does not handle NAs in the variables it uses to impute. So while clinchem_params + doseFlag + animal_no + timepoint + sacrifice + company was used for missForest, clinchem_params cannot be included in the CALIBERrfimputation. This is because the final imputation will be applied to data with real NAs in the clinchem_params. doseFlag + animal_no + timepoint + sacrifice + company do not have NAs. 

CALIBERrfimpute also does not handle categorical variables as input for imputation. 

Repeat the below process for MCAR, MAR and MNAR
```{r}
#CALIBERrfimpute
CALIBER<-list()
FULL<-list()
PARTIAL<-list()

imputedfull<-list()
imputedonly<-list()
#set up variables used in imputations, CALIBERrfimpute wont accept non-numericals
x<-SCdf[,impcols]
x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)


for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  
  #identify how many NAs in each collumn
idy<- length(which(is.na(t[[j]][,c(12)])))

#set up empty matrix of the correct size 
matrix_empty = matrix(nrow = idy, ncol = (end - start + 1))
emptyDF = data.frame(matrix_empty)
imp<- list()

for(i in start:ncol(t[[j]])){
  #imputation
  imp[[i]]<-CALIBERrfimpute::mice.impute.rfcont(y = t[[j]][,i], ry = !is.na(t[[j]][,i]), x = as.matrix(x))
  emptyDF[,i] <- imp[[i]]
}

#locate where NA's were created
idx<-as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))

#create full data frame containing the imputed values and partial data frame that has the original value and the imputed value side by side
imputedfull[[j]]<-t[[j]][,start:end]
idx$imputed<-idx$original<-NA
for(i in 1:nrow(idx)){
  idx$original[i]<-full[idx$row[i],idx$col[i]]
  idx$imputed<-unlist(imp[start:end])
  imputedonly[[j]]<-idx
  imputedfull[[j]][idx$row[i],idx$col[i]]<-as.numeric(idx$imputed[i])
}

new_list <- setNames(list(imputedfull[[j]]), placeHolder)
FULL <- append(FULL, new_list)  

new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
PARTIAL <- append(PARTIAL, new_list2) 

}
  
  CALIBER<- append(CALIBER, list(PARTIAL = PARTIAL,
                                 FULL = FULL))

#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(CALIBER[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(CALIBER[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(CALIBER[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- CALIBER[["PARTIAL"]][[j]][["original"]]
  xi <- CALIBER[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
CALIBER<- append(CALIBER, list(NRMSE = NRMSE,
                               partialMAE = partialMAE,
                               fullMAE = fullMAE))  
  
  
if((MCAR[1] %in% t[1]) ){
  CALIBER_MCAR <-CALIBER
}

if((MAR[1] %in% t[1]) ){
  CALIBER_MAR <-CALIBER
}

if((MNAR[1] %in% t[1]) ){
 CALIBER_MNAR <-CALIBER
}
  #CALIBER <- append(CALIBER$FULL, list(FULL = FULL))

# nrmse(ximp=as.matrix(imputedfull[[3]]), xmis=as.matrix(MAR[["missing 20 %"]][,12:22]), xtrue=as.matrix(SCdf[,12:22]))
# mae_MAR <- mean(abs(imputedonly[[1]][["original"]] - imputedonly[[1]][["imputed"]]))

```


##MICE
###MICE with clinchem
Repeat the below process for MCAR, MAR and MNAR
```{r}

#full<-SCdf[,12:22]
test<-list()
imp<-list()
imp2<-list()
imputedonly<-list()
MICErf<-list()
FULL<-list()
PARTIAL<-list()
for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  
  #make predictors numeric
  x<-SCdf[,impcols]
 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)
 
#make the predictor matrix using all covariates (including clin_chem parameters)
x<-cbind(x,t[[j]][,start:end])
init = mice(x, maxit=0) 
predM = init$predictorMatrix

#Do imputation
imp[[j]]<-mice(data = x, method="rf", predictorMatrix = predM)
imp2[[j]]<-merge_imputations(dat = x, imp = imp[[j]])
idx<- as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))
idx$imputed<-idx$original<-NA

for(i in 1:nrow(idx)){
  idx$original[i]<-full[idx$row[i],idx$col[i]]
  idx$imputed[i]<-imp2[[j]][idx$row[i],idx$col[i]]
  imputedonly[[j]]<-idx
}

new_list <- setNames(list(imp2[[j]]), placeHolder)
FULL <- append(FULL, new_list)  

new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
PARTIAL <- append(PARTIAL, new_list2) 
}

MICErf<- append(MICErf, list(PARTIAL = PARTIAL,
                             FULL = FULL))

#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(MICErf[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(MICErf[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(MICErf[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- MICErf[["PARTIAL"]][[j]][["original"]]
  xi <- MICErf[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
MICErf<- append(MICErf, list(NRMSE = NRMSE,
                             partialMAE = partialMAE,
                             fullMAE = fullMAE))


 if((MCAR[1] %in% t[1]) ){
  MICErf_MCAR <-MICErf
}

if((MAR[1] %in% t[1]) ){
  MICErf_MAR <-MICErf
}

if((MNAR[1] %in% t[1]) ){
  MICErf_MNAR <-MICErf
}

```

###MICE without clinchem
Repeat the below process for MCAR, MAR and MNAR
```{r}

#full<-SCdf[,12:22]
test<-list()
imp<-list()
imp2<-list()
imputedonly<-list()
ncMICErf<-list()
FULL<-list()
PARTIAL<-list()
for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  
  #make predictors numeric
x<-SCdf[,impcols]
  # x<-SCdf %>% 
  # select("Company", "Timepoint", "Animal_no.", "doseFlag", "Sacrifice_subgroup")
 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)
 
#make the predictor matrix using all covariates (including clin_chem parameters)
x<-cbind(x,t[[j]][,start:end])
init = mice(x, maxit=0) 
predM = init$predictorMatrix

#Do imputation
imp[[j]]<-mice(data = x, method="rf", predictorMatrix = predM)
imp2[[j]]<-merge_imputations(dat = x, imp = imp[[j]])
idx<- as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))
idx$imputed<-idx$original<-NA

for(i in 1:nrow(idx)){
  idx$original[i]<-full[idx$row[i],idx$col[i]]
  idx$imputed[i]<-imp2[[j]][idx$row[i],idx$col[i]]
  imputedonly[[j]]<-idx
}

new_list <- setNames(list(imp2[[j]]), placeHolder)
FULL <- append(FULL, new_list)  

new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
PARTIAL <- append(PARTIAL, new_list2) 
}

ncMICErf<- append(ncMICErf, list(PARTIAL = PARTIAL,
                                    FULL = FULL))

#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(ncMICErf[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(ncMICErf[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(ncMICErf[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- ncMICErf[["PARTIAL"]][[j]][["original"]]
  xi <- ncMICErf[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
ncMICErf<- append(ncMICErf, list(NRMSE = NRMSE,
                                 partialMAE = partialMAE,
                                 fullMAE = fullMAE))



 if((MCAR[1] %in% t[1]) ){
  ncMICErf_MCAR <-ncMICErf
}

if((MAR[1] %in% t[1]) ){
  ncMICErf_MAR <-ncMICErf
}

if((MNAR[1] %in% t[1]) ){
  ncMICErf_MNAR <-ncMICErf
}


```
clinchem_params + doseFlag + animal_no + timepoint + sacrifice + company
##missForest
###missForest with ClinChem
```{r}

FULL<-list()
PARTIAL <- list()
missForest<-list()
imp <- list()
imputedonly<-list()
nrmse<-list()
NRMSE<-list()
for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  placeHolder3<- paste0("NRMSE ", p[j] * 100, " % missing")
  
  #make predictors factors (can also be numeric, just not characters, max of 53 categories per variable allowed)
  x<-SCdf[,impcols]
 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 
 #set up data frame used in imputation 
 x <- cbind(x,t[[j]][,start:end])
 
 #impute
 imp[[j]] <- missForest(xmis = x, ntree = 10, variablewise = F)
 
 #locate where in the complete data the NAs were created
idx<- as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))

#make empty list to fill with original and imputed
idx$imputed<-idx$original<-NA

#start and end criteria 
start2 <- ncol(SCdf[,impcols]) +1
end2 <- ncol(x)

for(i in 1:nrow(idx)){
  
  #fill original column
  idx$original[i]<-full[idx$row[i],idx$col[i]]
  #fill imputed column
  idx$imputed[i]<-imp[[j]][["ximp"]][,start2:end2][idx$row[i],idx$col[i]]
  #store imputed only for each percentage of missingness
  imputedonly[[j]]<-idx
}

new_list <- setNames(list(imp[[j]][["ximp"]][,start2:end2]), placeHolder)
FULL <- append(FULL, new_list)  

 new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
 PARTIAL <- append(PARTIAL, new_list2) 

#extract NRMSE for each df 
nrmse[[j]]<-imp[[j]][["OOBerror"]][["NRMSE"]]
  new_list3 <- setNames(nrmse[[j]], placeHolder3)
 NRMSE <- append(NRMSE, new_list3) 

}

missForest<- append(missForest, list(PARTIAL = PARTIAL,
                                     FULL = FULL,
                                     NRMSE = NRMSE))

#fullMAE

fullmae<-list()
fullMAE<-list()
for(j in 1:length(missForest[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(missForest[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(missForest[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- missForest[["PARTIAL"]][[j]][["original"]]
  xi <- missForest[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
missForest<- append(missForest, list(partialMAE = partialMAE,
                                     fullMAE = fullMAE))  

#nameing

 if((MCAR[1] %in% t[1]) ){
  missForest_MCAR <-missForest
}

if((MAR[1] %in% t[1]) ){
  missForest_MAR <-missForest
}

if((MNAR[1] %in% t[1]) ){
  missForest_MNAR <-missForest
}

```
###missForest without ClinChem
```{r}

#missForest no clinchem
ncmissForest<-list()

FULL<-list()
PARTIAL<-list()
NRMSE<- list()
imputedfull<-list()
imputedonly<-list()
#full<-SCdf[,12:22]
NRMSEncmissForest<-list()

for(j in 1:length(t)){
  placeHolder <- paste0("FULLimputed ", p[j] * 100, " %")
  placeHolder2 <- paste0("imputed vs full ", p[j] * 100, " %")
  
  #identify how many NAs in each collumn
idy<- length((t[[1]][,start]))

#set up empty matrix of the correct size 
matrix_empty = matrix(nrow = idy, ncol = 11)
emptyDF = data.frame(matrix_empty)
imp<- list()

for(i in start:ncol(t[[j]])){
  x<-SCdf[,impcols] 
  
  start2 <- ncol(SCdf[,impcols]) +1
 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 #x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)
 
 x <- cbind(x,t[[j]][,i])
  
  #imputation, makes full data frame, but imputation done without clinchem
  imp[[i]] <- missForest(xmis = x, ntree = 10, variablewise = F)
  emptyDF[,i] <- imp[[i]][["ximp"]][[start2]]
  imputedfull[[j]]<- emptyDF
}

#locate where NA's were created
idx<-as.data.frame(which(is.na(t[[j]][,start:end]), arr.ind=TRUE))

#create full data frame containing the imputed values and partial data frame that has the original value and the imputed value side by side

idx$imputed<-idx$original<-NA
for(i in 1:nrow(idx)){
  idx$original[i]<-full[idx$row[i],idx$col[i]]
  idx$imputed[i]<-imputedfull[[j]][,start:end][idx$row[i],idx$col[i]]
  imputedonly[[j]]<-idx
}
new_list <- setNames(list(imputedfull[[j]][,start:end]), placeHolder)
FULL <- append(FULL, new_list)  

new_list2 <- setNames(list(imputedonly[[j]]), placeHolder2)
PARTIAL <- append(PARTIAL, new_list2) 

}
  
 ncmissForest<- append(ncmissForest, list(PARTIAL = PARTIAL,
                                          FULL = FULL))
 
#NRMSE + fullMAE
nrmse <-list()
fullmae<-list()
NRMSE<-list()
fullMAE<-list()
for(j in 1:length(ncmissForest[["FULL"]])){
  
placeHolder3 <- paste0( p[,j] * 100, " % missing")  

xi<- as.matrix(ncmissForest[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(t[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi, xmis = xm, xtrue = xt)

new_list <- setNames((nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 

fullmae[j] <- mean(abs(xt - xi))

new_list2 <- setNames((fullmae[j]), placeHolder3)
fullMAE <- append(fullMAE, new_list2) 

}
 
#partialMAE
 partialmae <- list()  
 partialMAE<-list()
  for(j in 1:length(ncmissForest[["PARTIAL"]])){
  placeHolder3 <- paste0( p[,j] * 100, " % missing")  
    
  xt <- ncmissForest[["PARTIAL"]][[j]][["original"]]
  xi <- ncmissForest[["PARTIAL"]][[j]][["imputed"]]
  
  partialmae[j] <- mean(abs(xt - xi))
  
  new_list <- setNames((partialmae[j]), placeHolder3)
  partialMAE <- append(partialMAE, new_list) 
  }
 
ncmissForest<- append(ncmissForest, list(NRMSE = NRMSE,
                                         partialMAE = partialMAE,
                                         fullMAE = fullMAE))  
 
 if((MCAR[1] %in% t[1]) ){
  ncmissForest_MCAR <-ncmissForest
}

if((MAR[1] %in% t[1]) ){
  ncmissForest_MAR <-ncmissForest
}

if((MNAR[1] %in% t[1]) ){
  ncmissForest_MNAR <-ncmissForest
}

```


#METRICS
```{r}
# MICErf[["NRMSE"]]
# missForest[["NRMSE"]]
# ncMICErf[["NRMSE"]]
# ncmissForest[["NRMSE"]]
df<- as.data.frame(rbind(CALIBER[["NRMSE"]], 
                         ncMICErf[["NRMSE"]],
                         ncmissForest[["NRMSE"]],
                         MICErf[["NRMSE"]],
                         missForest[["NRMSE"]]))
rownames(df) <- c("ncCALIBERrfimoute","ncMICErf","ncmissForest", "MICErf", "missForest")

ggplot(data= df, aes(x= ))

```



#metrics practice
###NRMSE
```{r}
nrmse <-list()
mae<-list()
NRMSE<-list()
for(j in 1:length(ncmissForest_MAR[["FULL"]])){
  
placeHolder3 <- paste0("NRMSE ", p[,j] * 100, " % missing")  

xi<- as.matrix(ncmissForest_MAR[["FULL"]][[j]])
xt<-as.matrix(SCdf[,start:end])
xm<-as.matrix(MAR[[j]][,start:end]) 
  
nrmse[[j]]<-missForest::nrmse(ximp = xi,xmis = xm, xtrue = xt)
mae[[j]] <- mean(abs(xt - xi))

new_list <- setNames(list(nrmse[j]), placeHolder3)
NRMSE <- append(NRMSE, new_list) 
}


```

###MAE
```{r}
# mae_MCAR <- mean(abs(orig_MCAR - imp_MCAR))
#   mae_MAR <- mean(abs(orig_MAR - imp_MAR))
#   mae_MNAR <- mean(abs(orig_MNAR - imp_MNAR))

mae <- list()  
  for(j in 1:length(ncmissForest_MAR[["PARTIAL"]])){
  xt <- ncmissForest_MAR[["PARTIAL"]][[j]][["original"]]
  xi <- ncmissForest_MAR[["PARTIAL"]][[j]][["imputed"]]
  
  mae[[j]] <- mean(abs(xt - xi))
  }

  
```

#IMPUTATIONtrials
#####MICE trial
```{r}
#rf MICE
library(mice)

# mice::mice.impute.rf()
# merge_imputations(dat = SCdf5, imp = imp)

#still need to convert to numeric 
x<-SCdf %>% 
  select("Company", "Toxin", "Animal_no.", "doseFlag", "Sacrifice_subgroup")
 x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
 x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)
# full<-SCdf[,12:22]
x<-cbind(x,MAR[[1]][,12:22])

#using all covariates (including clin_chem parameters)
init = mice(x, maxit=0) 
predM = init$predictorMatrix
imp <- mice(data = x, method="rf", predictorMatrix = predM)

#not using clin_chem parameters

predM[6:16,6:16]<-0
imp <- mice(data = x, method="rf", predictorMatrix = predM)

#merges multiple data frames/picks optinal imputation, returns full dataframe with imputed values
imp2<-merge_imputations(dat = x, imp = imp)


#imp<- mice::mice.impute.rf(y = MNAR[[1]][,12], ry = !is.na(MNAR[[1]][,12]), x = as.matrix(x))
```
####CALIBERrfimputeattempt

```{r}
#CALIBERrfimpute

imp<-list()
x<-SCdf %>% 
  select("Company", "Toxin", "Animal_no.", "doseFlag", "Sacrifice_subgroup")
x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)

idx<- length(which(is.na(MAR[["missing 5 %"]][,12])))
matrix_empty = matrix(nrow = idx, ncol = 11)
test = data.frame(matrix_empty)

for(i in 12:ncol(MAR[["missing 5 %"]])){
  #imputation
  imp[[i]]<-CALIBERrfimpute::mice.impute.rfcont(y = MAR[["missing 5 %"]][,i], ry = !is.na(MAR[["missing 5 %"]][,i]), x = as.matrix(x))
  test[,i] <- imp[[i]]
}






imp<-CALIBERrfimpute::mice.impute.rfcont(y = MAR[["missing 5 %"]][,12], ry = !is.na(MAR[["missing 5 %"]][,12]), x = as.matrix(x))
matrix_empty = matrix(nrow = idx, ncol = 1)
test = data.frame(matrix_empty)
test[,2] <- imp



CALIBERrfimpute::mice.impute.rfcont(y = MARSdf05$S_Creatinine, ry = !is.na(MARSdf05$S_Creatinine), x = as.matrix(MARSdf05[,1:9]))

mice.impute.rfcont5 <- function(y, ry, x, ...){
CALIBERrfimpute::mice.impute.rfcont(
y = y, ry = ry, x = x, ntree_cont = 1)
}

mice.impute.rfcont5(y = as.numeric(SCdf5$S_Creatinine), ry = !is.na(SCdf5$S_Creatinine), x = SCdf5$Animal_no.)


```
####IMPUTATION CALIBER PRACTICE 
CALIBERrfimpute also does not handle categorical variables as input for imputation. 
```{r}
#CALIBERrfimpute

imp<-list()

#set up variables used in imputations, CALIBERrfimpute wont accept non-numericals
x<-SCdf %>% 
  select("Company", "Toxin", "Animal_no.", "doseFlag", "Sacrifice_subgroup")
x[sapply(x, is.character)] <- lapply(x[sapply(x, is.character)], as.factor)
x[sapply(x, is.factor)] <- lapply(x[sapply(x, is.factor)], as.numeric)

#identify how many NAs in each collumn
idx<- length(which(is.na(MAR[["missing 5 %"]][,12])))

#set up empty matrix of the correct size 
matrix_empty = matrix(nrow = idx, ncol = 11)
test = data.frame(matrix_empty)

for(i in 12:ncol(MAR[["missing 5 %"]])){
  #imputation
  imp[[i]]<-CALIBERrfimpute::mice.impute.rfcont(y = MAR[["missing 5 %"]][,i], ry = !is.na(MAR[["missing 5 %"]][,i]), x = as.matrix(x))
  test[,i] <- imp[[i]]
}
```
full df with imputed
```{r}
# SCdf[idx$row[1],idx$col[1]]
# 
test2[23,1]
# 29.172
test2<-MAR[["missing 5 %"]][,12:22]
for(i in 1:nrow(idx)){
 test2[idx$row[i],idx$col[i]]<-as.numeric(idx$imputed[i])
}
```
side by side original and imputed
```{r}
idx<-as.data.frame(which(is.na(MAR[["missing 5 %"]][,12:22]), arr.ind=TRUE))
test<-SCdf[,12:22]
test2<-MAR[["missing 5 %"]][,12:22]
idx$imputed<-idx$original<-NA
for(i in 1:nrow(idx)){
  idx$original[i]<-test[idx$row[i],idx$col[i]]
  idx$imputed<-unlist(imp[12:22])
  test2[idx$row[i],idx$col[i]]<-as.numeric(idx$imputed[i])
  #idx$imputed[i]<-MAR[["missing 5 %"]][idx$row[i],idx$col[i]]
}
```

#Density (violin + boxplot)
#External check
```{r}
#find relationship to test bias. Serum Albumin and Total Protein have a strong relationship. 
#df2 <- df[rowSums(is.na(df[,15:30])) > 0, ]  
firstcol = which(colnames(SCdf)=="S_Creatinine")
lastcol = which(colnames(SCdf)=="S_Total_protein")

pairs(SCdf[c(firstcol:lastcol)])

ggplot(data = SCdf, aes(x = S_Total_protein, y = S_Albumin)) +
  geom_point() +
   geom_smooth(method = lm,
               level = 0.95) +
  stat_poly_eq(data=SCdf, aes(x=S_Total_protein, y = S_Albumin))
  

ggplot(data = SCdf, aes(x = S_ALT, y = S_AST)) +
  geom_point() +
 geom_smooth(method = lm,
               level = 0.95) +
  stat_poly_eq(data=SCdf, aes(x=S_ALT, y = S_AST))


ggpubr::stat_cor(method = "pearson",
                   hjust = 0,
                   vjust = 0) 
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
